---
title: 'Analysing Word Document frequency : Tf-Idf'
author: "Saumya Mehta"
date: "5/14/2018"
output: html_document
fig_caption: yes
---

<style type="text/css">
.caption {
    font-size: 16pt;
    font-family: raleway;
    color : #9E9E9E;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Term frequency can be considered as one measure of how important a word is. It indicates how frequently a word occurs in a document. <br>
Inverse Documnent Frequency increases weight for rare words. Tf-idf measures importance of rare words.<br>
Idf is given by the following equation :<br>
<math xmlns="http://www.w3.org/1998/Math/MathML" display="block">
  <mi>i</mi>
  <mi>d</mi>
  <mi>f</mi>
  <mo stretchy="false">(</mo>
  <mtext>term</mtext>
  <mo stretchy="false">)</mo>
  <mo>=</mo>
  <mi>ln</mi>
  <mo>&#x2061;<!-- a --></mo>
  <mrow class="MJX-TeXAtom-ORD">
    <mrow>
      <mo>(</mo>
      <mfrac>
        <msub>
          <mi>n</mi>
          <mrow class="MJX-TeXAtom-ORD">
            <mtext>documents</mtext>
          </mrow>
        </msub>
        <msub>
          <mi>n</mi>
          <mrow class="MJX-TeXAtom-ORD">
            <mtext>documents containing term</mtext>
          </mrow>
        </msub>
      </mfrac>
      <mo>)</mo>
    </mrow>
  </mrow>
</math>


## Term frequency in Jane Austen's novel.
Finding the most common words in novels

```{r}
library(dplyr)
library(janeaustenr)
library(tidytext)

book_words <- austen_books() %>% 
  unnest_tokens(word,text) %>% 
  count(book, word, sort = T) %>% 
  ungroup()

total_words <- book_words %>% 
  group_by(book) %>% 
  summarise(total =sum(n))


book_words <- left_join(book_words,total_words)
book_words
```


n = number of times a word was used in the book. 
total = total number of words in the book.

```{r figs1, echo=FALSE,fig.cap="\\label{fig:figs1}Figure 3.1: Term Frequency Distribution in Jane Austen’s Novels"}

library(ggplot2)

ggplot(book_words, aes(n/total, fill = book)) +
  geom_histogram(show.legend = F) +
  xlim(NA, 0.0009) +
  facet_wrap(~book, ncol = 2, scales = "free_y")+theme_light()
```


## Zipf's Law
Frequency of the word is inversely proportional to it's rank.

```{r}
freq_by_rank <- book_words %>% 
  group_by(book) %>% 
  mutate(rank = row_number(),
          `term frequency` = n/total) 
```

Zipf's Law visualization

```{r figs2, echo=FALSE,fig.cap="\\label{fig:figs2}Figure 3.2: Zipf’s law for Jane Austen’s novels"}

freq_by_rank %>% 
  ggplot(aes(rank, `term frequency`, color = book)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = F) + 
  scale_x_log10() + 
  scale_y_log10()+theme_light()
```


## Bind tf-idf function
bind_tf_idf takes as input tidy dataset with one row per token per document form. One column contains the token and one column contains the document. For bind_tf_idf function, we only need all the terms in the document.

```{r}
book_words <- book_words %>% 
  bind_tf_idf(word, book, n)
book_words
```

idf and hence tf-idf is zero for most common words. idf is nearly zero for terms that occur in many documents in the collection. 

```{r}
book_words %>% 
  select(-total) %>% 
  arrange(desc(tf-idf))
```

```{r figs3, echo=FALSE,fig.cap="\\label{fig:figs3}Figure 3.3: Highest tf-idf words in each of Jane Austen’s Novels"}
book_words %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(book) %>% 
  top_n(15) %>% 
  ungroup() %>% 
  ggplot(aes(word, tf_idf, fill = book)) + 
  geom_col(show.legend = F) +
  labs(
    x = NULL,
    y = "tf-idf"
  ) +
  facet_wrap(~book, ncol = 2, scales = "free") + 
  coord_flip() +
  theme_light()
```

tf-idf indentifies the words that are important to a document in the collection of documents.


## Corpus of physical texts 
Project Gutenberg and apply tf-idf on the corpus.

```{r}
library(gutenbergr)

physics <- gutenberg_download(c(37729, 14725, 13746, 5001),
                              meta_fields = "author")
```

Now after extracting the texts, we need to  unnest into tokens and count the occurence of each words.

```{r}
physics_words <- physics %>%
  unnest_tokens(word, text) %>%
  count(author, word, sort = TRUE) %>%
  ungroup()

physics_words
```

Calculating tf-idf and visualizing high tf-idf words

```{r figs4, echo=FALSE,fig.cap="\\label{fig:figs4}Figure 3.4: Highest tf-idf words in each physics texts"}
plot_physics <- physics_words %>%
  bind_tf_idf(word, author, n) %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>%
  mutate(author = factor(author, levels = c("Galilei, Galileo",
                                            "Huygens, Christiaan", 
                                            "Elliot, Robert H. (Robert Henry)",
                                            "Einstein, Albert")))

plot_physics %>% 
  group_by(author) %>% 
  top_n(15, tf_idf) %>% 
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(word, tf_idf, fill = author)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~author, ncol = 2, scales = "free") +
  coord_flip() + 
  theme_light()
```

```{r}
library(stringr)

physics %>% 
  filter(str_detect(text, "eq\\.")) %>% 
  select(text)
```

```{r figs5, echo=FALSE,fig.cap="\\label{fig:figs5}Figure 3.5: Highest tf-idf words in classic physics texts"}
mystopwords <- data_frame(word = c("eq", "co", "rc", "ac", "ak", "bn", 
                                   "fig", "file", "cg", "cb", "cm"))
physics_words <- anti_join(physics_words, mystopwords, by = "word")
plot_physics <- physics_words %>%
  bind_tf_idf(word, author, n) %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>%
  group_by(author) %>% 
  top_n(15, tf_idf) %>%
  ungroup %>%
  mutate(author = factor(author, levels = c("Galilei, Galileo",
                                            "Huygens, Christiaan",
                                            "Elliot, Robert H. (Robert Henry)",
                                            "Einstein, Albert")))

ggplot(plot_physics, aes(word, tf_idf, fill = author)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~author, ncol = 2, scales = "free") +
  coord_flip() + 
  theme_light()
```

