---
title: "Juggling between tidy and non tidy forms"
author: "Saumya Mehta"
date: "5/15/2018"
output: html_document
fig_caption: yes
---
<style type="text/css">
.caption {
    font-size: 16pt;
    font-family: raleway;
    color : #9E9E9E;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidying a document term matrix 
* Each row represents one document
* Each column represents one term
* Each value (typically) contains the number of appearances of that term in that document. <br>
DTMs are usually sparse matrices. DTMs cannot be used directly with tidy tools. Two Packages to convert between two formats :<br>
      1.) tidy() : dtm to tidy data <br>
      2.) cast() : tidy data to dtm (cast_sparse(), cast_dtm(), cast_dfm())

## Tidying dtm object
DocumentTermMatrix class from tm packages. 
```{r}
library(tm)

data("AssociatedPress", package = "topicmodels")

AssociatedPress
```

Access the terms in a dtm using Terms() function.
```{r}
terms <- Terms(AssociatedPress)
head(terms)
```

To analyse with tidy tools, we need to convert it into tidy data using tidy(dtm object). 
```{r}
library(dplyr)
library(tidytext)

ap_td <- tidy(AssociatedPress)

ap_td
```

Tidied version has no rows where count = 0. This form is convinient for analysis with dplyr, tidytext and ggplot2. 

```{r}
ap_sentiments <- ap_td %>% 
  inner_join(get_sentiments("bing"), by = c(term = "word"))

ap_sentiments
```

Analysing contribution to the sentiment.
```{r figs, echo=FALSE,fig.cap="\\label{fig:figs}Figure 5.1: Words from AP articles with the greatest contribution to positive or negative sentiments, using the Bing sentiment lexicon"}
library(ggplot2)

ap_sentiments %>% 
  count(sentiment, term, wt = count) %>% 
  ungroup() %>% 
  filter(n>=200) %>% 
  mutate(n = ifelse(sentiment == "negative", -n,n)) %>% 
  mutate(term = reorder(term,n)) %>% 
  ggplot(aes(term,n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  ylab ("Contribution to the sentiment") +
  coord_flip() +
  theme_light()

```


## Tidying dfm objects
Document Feature Matrix class from quanteda package. 

```{r}
library(quanteda)
data("data_corpus_inaugural", package = "quanteda")
inaug_dfm <- dfm(data_corpus_inaugural, verbose = F)

inaug_tidy <- tidy(inaug_dfm)

inaug_tidy
```

We can find the words most specific to each inaugural speeches using tf-idf mapping.

```{r}
inaug_tf_idf <- inaug_tidy %>% 
  bind_tf_idf(term,document,count) %>% 
  arrange(desc(tf_idf))

inaug_tf_idf
```

tidyr's complete() function is used to include zeros where the word didn't appear in the document.
```{r}
library(tidyr)

year_term_counts <- inaug_tidy %>% 
  extract(document, "year","(\\d+)",convert = T) %>% 
  complete(year, term, fill = list(count = 0)) %>% 
  group_by(year) %>% 
  mutate(year_total = sum(count))
```

Here we pick up a few words and check how they changed in frequency over time.
```{r figs1, echo=FALSE,fig.cap="\\label{fig:figs1}Figure 5.2: Changes in word frequency over time within Presidential inaugural addresses, for four selected terms"}
year_term_counts %>% 
  filter(term %in% c("god","america","foreign", "union","constitution","freedom")) %>% 
  ggplot(aes(year, count/year_total)) +
  geom_point() + 
  geom_smooth() + 
  facet_wrap(~term, scales = "free") + 
  scale_y_continuous(labels = scales::percent_format()) +
  ylab("% frequency of words in inaugural address")
  
```


## Casting tidytext data into a matrix

```{r}
ap_td %>%
  cast_dtm(document, term, count)

ap_td %>%
  cast_dfm(document, term, count)

library(Matrix)

# cast into a Matrix object
m <- ap_td %>%
  cast_sparse(document, term, count)

class(m)

dim(m)

```

```{r}
library(janeaustenr)

austen_dtm <- austen_books() %>%
  unnest_tokens(word, text) %>%
  count(book, word) %>%
  cast_dtm(book, word, n)

austen_dtm
```


## Tidying corpus objects with metadata
A corpus object is structured like a list each containing both text and metadata. This is a flexible method os storage for documents but doesn't lend itself to processing with tidy tools. <br>
We can thus use the tidy() method to construct a table with one row per document, including the metadata (such as id and datetimestamp) as columns alongside the text



